Accuracy and precision of zero-heat-flux temperature monitoring

### 2019-11-27 manuscript/manuscript.Rmd
added beth and edits from KC

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring"
description: A systematic review and meta-analysis
author: 
- name: Aaron Conway
  affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: aaron.conway@utoronto.ca
  url: https://www.aaronconway.info
- name: Megan Bittner
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Dan Phan
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Navpreet Kamboj
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: navpreet.kamboj@mail.utoronto.ca
- name: Kristina Chang
  affiliation: Toronto General Hospital, UHN
- name: Peter Collins
  affiliation: Toronto General Hospital, UHN
- name: Matteo Parotto
  affiliation: Toronto General Hospital, UHN
date: "`r Sys.Date()`"
bibliography: bibliography.bib
output: distill::distill_article
repository_url: https://github.com/awconway/zhf-review
twitter:
  creator: "@aw_conway"
---

@@

Observational studies that reported temperature measurements from a zero-heat-flux thermometer and comparator thermometer were included. Studies involving a case control design were excluded due to potential for overestimation of the intervention performance. Studies were excluded if conducted on non-human subjects or outside of a clinical healthcare setting. No publication date restrictions were applied.Published conference abstracts were included if there was enough information reported to appraise the quality of the study. There were no language restrictions applied during the search.

@@

Published studies were found by searching Medline and EMBASE from January 2000 to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located [here](https://aconway.dev/zhf app). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

Two reviewers independently assessed the risk of bias for the included studies using the revised Quality Assessment of Diagnostic Accuracy Studies (QUADAS-2).[@whiting2011quadas] Reviewers rated the risk of bias for patient selection, conduct of the zero-heat-flux measurements, conduct of the comparator thermometer measurements, and timing and flow (eg, timing of ZHF and established core temperature measurements, dropouts) as “high”, “low” or “unclear” risk of bias. We worked to minimize the risk of publication bias by conducting a comprehensive search of multiple databases as well as an international clinical trial registry.[@glasziou2001systematic] Statistical approaches for detection of reporting bias were not conducted due to lack of validated methods.[@begg2005systematic] Simulations have revealed that tests for detecting funnel plot asymmetry will result in publication bias being incorrectly identified too often.[@deeks2005performance]

@@

The objective for the meta-analysis was to estimate the population limits of agreement between temperature measurements from the ZHF and established comparator thermometers. A framework for meta-analysis of Bland-Altman method comparison studies developed by @tipton2017framework based on limits of agreement approach was used. This method was selected because it parallels the approach used in primary Bland-Altman studies, whereby an estimate is generated for the pooled limits of agreement in the population (not just in the samples studied). The ‘population limits of agreement’ is more broad than the limits of agreement commonly reported in the meta-analyses of Bland-Altman studies.[@tipton2017framework] Pooled limits of agreement were calculated using $\delta\pm2\sqrt{\sigma2+\tau2}$, where $\delta$ is the average bias across studies, $\sigma2$ is the average within-study variation in differences and $\tau2$ is the variation in bias across studies.

Estimations of $\delta$ and $\sigma2$ were made using a weighted least-squares model (similar to a random-effects approach), with the associated standard errors estimated using robust variance estimation. Robust variance estimation was used alternatively to model-based standard errors as some studies included in the systematic review used repeated-measures designs without accommodating for the correlation between measurements[@hedges2010robust; @tanner2016handling; @tipton2015small] We used the method-of-moments estimator from @dersimonian1986meta for the $\tau2$ parameter.

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses[@team2017r] All data and R code used in the meta-analyses can be located [here](https://github.com/awconway/zhf-review).

Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator thermometer minus ZHF thermometer measured in °C. In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'. Other studies reported multiple sets of results, whereby analyses were conducted between ZHF and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were apart of separate meta-analyses groups. One study reported intraoperative, postoperative and overall results for the same participants. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups.

The conventionally cited clinically acceptable agreement between ZHF and comparator devices is 0.5 °C.[@eshraghi2014] It was deemed that outer confidence bounds for 95% limits of agreement between zero-heat-flux and core temperature measurements (termed as ‘population limits of agreement’) outside of these bounds would be clinically unacceptable.

A sensitivity analysis for the primary comparison (ZHF versus temperature measurement at core site) was performed based on risk of bias, whereby ‘unclear risk of bias’ was treated as ‘high risk’ and ‘high risk of bias’ studies were excluded from analyses. We also conducted a sensitivity analysis for the primary comparison (ZHF versus temperature measurement at core site) excluding studies that received funding from industry. As clinicians would be interested in the accuracy of ZHF relative within the clinical setting in which they use it, we conducted subgroup analyses for the primary comparison according to clinical setting of the study (either intraoperative or intensive care unit).

@@

All studies included in this systematic review evaluated the zero-heat-flux temperature monitoring system manufactured by 3M. Previously known as the SpotOn Temperature Monitoring System, the 3M ZHF device is now referred to commercially as the Bair Hugger Temperature Monitoring System. All comparisons reported adherence to the zero-heat-flux device manufacturer instructions and placed the sensor on the forehead of the participants. One included study also reported results for a comparison where the zero-heat-flux thermometer was placed on the forehead. We did not included this comparison in our meta-analysis.

High risk of bias was associated with patient selection for 13 (43%) comparisons from 9 (56%) studies, conduct of ZHF and comparator measurements in 10(33%) comparisons from 8 (50%) studies and 12 (40%) comparisons 9 (30%) studies, respectively (mostly due to ZHF measurements being taken with knowledge of the comparator measurements and vice versa) and participant flow for 13 (43%) studies. In 19 (50%) comparisons from 8 (50%) studies, the authors had declared a conflict of interest or receipt of funding or equipment from the manufacturer of the ZHF device under evaluation.

@@

Although comparisons with core temperature are the most important to evaluate, we included studies that compared zero-heat-flux and peripheral temperature monitoring devices in our review because clinicians may be interested to determine if this new technology is sufficiently accurate to replace other ways non-invasive monitor temperature. Also, there may be situations where it is not practical to expose research participants to the potential additional risk of invasive core temperature monitoring for the purposes of evaluating the new temperature monitoring device when it is standard practice for only peripheral temperature monitoring to be used. Whether or not the zero-heat-flux thermometer is sufficiently accurate to be used in place of nasopharyngeal thermometers is unclear. Population limits of agreement calculated from four studies were broad, spanning from one degree less than zero-heat-flux temperature to one degree higher. Further studies are therefore required.

@@

Many studies in this review analysed a large number of measurements of temperature with relatively small sample sizes. Importantly, the approach we used for our meta-analysis takes this into account. By using robust variance estimation, weights for pooling estimates in the meta-analysis become proportional to the number of patients not the total number of measurements.[@tipton2017framework]

>>>>

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring"
description: A systematic review and meta-analysis
author: 
- name: Aaron Conway
  affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: aaron.conway@utoronto.ca
  url: https://www.aaronconway.info
- name: Megan Bittner
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Dan Phan
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Navpreet Kamboj
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: navpreet.kamboj@mail.utoronto.ca
- name: Kristina Chang
  affiliation: Toronto General Hospital, UHN
- name: Peter Collins
  affiliation: Toronto General Hospital, UHN
- name: Elizabeth Tipton
  affiliation: Dapartment of Statistics, Northwestern University
- name: Matteo Parotto
  affiliation: Toronto General Hospital, UHN
date: "`r Sys.Date()`"
bibliography: bibliography.bib
output: distill::distill_article
repository_url: https://github.com/awconway/zhf-review
twitter:
  creator: "@aw_conway"
---

@@

Observational studies that reported temperature measurements from a zero-heat-flux thermometer and comparator thermometer were included. Studies involving a case control design were excluded due to potential for overestimation of the intervention performance. Studies were excluded if conducted on non-human subjects or outside of a clinical healthcare setting. No publication date restrictions were applied. Published conference abstracts were included if there was enough information reported to appraise the quality of the study. There were no language restrictions applied during the search.

@@

Published studies were found by searching Medline and EMBASE to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located [here](https://aconway.dev/zhf app). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

Two reviewers independently assessed the risk of bias for the included studies using the revised Quality Assessment of Diagnostic Accuracy Studies (QUADAS-2).[@whiting2011quadas] Reviewers rated the risk of bias for patient selection, conduct of the zero-heat-flux measurements, conduct of the comparator thermometer measurements, and timing and flow (eg, timing of zero-heat-flux and established core temperature measurements, dropouts) as “high”, “low” or “unclear” risk of bias. We worked to minimize the risk of publication bias by conducting a comprehensive search of multiple databases as well as an international clinical trial registry.[@glasziou2001systematic] Statistical approaches for detection of reporting bias were not conducted due to lack of validated methods.[@begg2005systematic] Simulations have revealed that tests for detecting funnel plot asymmetry will result in publication bias being incorrectly identified too often.[@deeks2005performance]

@@

The objective for the meta-analysis was to estimate the population limits of agreement between temperature measurements from the zero-heat-flux and established comparator thermometers. A framework for meta-analysis of Bland-Altman method comparison studies based on limits of agreement approach was used.[@tipton2017framework]  This method was selected because it parallels the approach used in primary Bland-Altman studies, whereby an estimate is generated for the pooled limits of agreement in the population (not just in the samples studied). The ‘population limits of agreement’ is more broad than the limits of agreement commonly reported in the meta-analyses of Bland-Altman studies.[@tipton2017framework] Pooled limits of agreement were calculated using $\delta\pm2\sqrt{\sigma2+\tau2}$, where $\delta$ is the average bias across studies, $\sigma2$ is the average within-study variation in differences and $\tau2$ is the variation in bias across studies.

Estimations of $\delta$ and $\sigma2$ were made using a weighted least-squares model (similar to a random-effects approach), with the associated standard errors estimated using robust variance estimation. Robust variance estimation was used alternatively to model-based standard errors as some studies included in the systematic review used repeated-measures designs without accommodating for the correlation between measurements.[@hedges2010robust; @tanner2016handling; @tipton2015small] We used the method-of-moments estimator from DerSimonian & Laird @dersimonian1986meta for the $\tau2$ parameter.

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses[@team2017r] All data and R code used in the meta-analyses can be located [here](https://github.com/awconway/zero-heat-flux-review).

Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator thermometer minus zero-heat-flux thermometer measured in °C. In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'. Other studies reported multiple sets of results, whereby analyses were conducted between zero-heat-flux and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were a part of separate meta-analysis groups. One study reported intraoperative, postoperative and overall results for the same participants. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups.

The conventionally cited clinically acceptable agreement between zero-heat-flux and comparator devices is 0.5 °C.[@eshraghi2014] It was deemed that outer confidence bounds for 95% limits of agreement between zero-heat-flux and core temperature measurements (termed as ‘population limits of agreement’) outside of these bounds would be clinically unacceptable.

A sensitivity analysis for the primary comparison (zero-heat-flux versus temperature measurement at core site) was performed based on risk of bias, whereby ‘unclear risk of bias’ was treated as ‘high risk’ and ‘high risk of bias’ studies were excluded from analyses. We also conducted a sensitivity analysis for the primary comparison (zero-heat-flux versus temperature measurement at core site) excluding studies that received funding from industry. As clinicians would be interested in the accuracy of zero-heat-flux relative within the clinical setting in which they use it, we conducted subgroup analyses for the primary comparison according to clinical setting of the study (either intraoperative or intensive care unit).

@@

All studies included in this systematic review evaluated the zero-heat-flux temperature monitoring system manufactured by 3M. Previously known as the SpotOn Temperature Monitoring System, the 3M zero-heat-flux device is now referred to commercially as the Bair Hugger Temperature Monitoring System. All comparisons reported adherence to the zero-heat-flux device manufacturer instructions and placed the sensor on the forehead of the participants. One included study also reported results for comparisons where the zero-heat-flux thermometer was not placed on the forehead. We did not include these comparisons in our meta-analysis.

High risk of bias was associated with patient selection for 13 (43%) comparisons from 9 (56%) studies, conduct of zero-heat-flux and comparator measurements in 10 (33%) comparisons from 8 (50%) studies and 12 (40%) comparisons 9 (30%) studies, respectively (mostly due to zero-heat-flux measurements being taken with knowledge of the comparator measurements and vice versa) and participant flow for 13 (43%) studies. In 19 (50%) comparisons from 8 (50%) studies, the authors had declared a conflict of interest or receipt of funding or equipment from the manufacturer of the zero-heat-flux device under evaluation.

@@

Although comparisons with core temperature are the most important to evaluate, we included studies that compared zero-heat-flux and peripheral temperature monitoring devices in our review because clinicians may be interested to determine if this new technology is sufficiently accurate to replace other ways to non-invasively monitor temperature. Also, there may be situations where it is not practical to expose research participants to the potential additional risk of invasive core temperature monitoring for the purposes of evaluating the new temperature monitoring device when it is standard practice for only peripheral temperature monitoring to be used. Whether or not the zero-heat-flux thermometer is sufficiently accurate to be used in place of nasopharyngeal thermometers is unclear. Population limits of agreement calculated from four studies were broad, spanning from one degree less than zero-heat-flux temperature to one degree higher. Further studies are therefore required.

@@

Many studies in this review analysed a large number of measurements of temperature with relatively small sample sizes. Importantly, the approach we used for our meta-analysis takes this into account. By using robust variance estimation, weights for pooling estimates in the meta-analysis become proportional to the number of patients, not the total number of measurements.[@tipton2017framework]
